Generative AI for image reconstruction in Intensity Interferometry: a first attempt

### 2025-12-07 Phase_II/draft/ParaGan.tex
to Prof. Saha

In the training of GANs, one often-followed strategy to potentially boost performance is to give the Discriminator an advantage by increasing its number of training epochs before returning to the Generator's training. This hyperparameter is referred to as the Discriminator repetition (as seen in Table \ref{tab:hyperparameters}. While this can lower the Discriminator loss, as shown in Fig.~\ref{fig:Plot_discrep_loss}, it also increases training time. In training our model, we did not notice any significant advantage derived from this strategy. Since the generated images did not noticeably improve with additional Discriminator training, we adopted the strategy of training both the networks with equal preference (Discriminator repetition = 1).

@@

An optimized set of hyperparameters is selected through an iterative process of training and validation. For a tentatively chosen set of the hyperparameters, the model is trained using the Training Set until the both the Discriminator and the Generator loss functions are minimized. This model, thus trained along with its model parameters, is then passed through validation using the Validation Set. This cycle of training and validation is iterated till an optimal set of hypermaraters is arrived at. During each epoch of training of the model, both the Discriminator and the Generator networks are trained for 100,000 steps. Plots of the ''Discriminator Loss'' function and the ``Generator Total Loss'' function presented in Fig.{\ref{fig:Plot_discrep_loss}} and Fig.{\ref{fig:Plot_telescopes_loss} represent the results of this training for two of the hyperparameters, namely, the Disriminator repetition (``discrep'', in short) and the Number of telescopes, respectively}. Obviously, the most compute-intensive part of this process is that of the training of the Model. The results of training and validation presented in this work were carried out on a CPU using two nodes, each with 96 threads and the entire process of training and validation required approximately 20 hours on the machine employed for this work. This iterative process of training and validation of the Model is represented schematically in the Fig.\ref{fig:FlowDiagram}. The chosen optimal set of hyperparameters is presented in the table Table-{\ref{tab:hyperparameters}.}
\begin{table}[ht]
	\centering
	\caption{Selected hyperparameters used for training the model.}
	\label{tab:hyperparameters}
	\begin{tabular}{ll}
		\hline
		\textbf{Hyperparameter} & \textbf{Value} \\
		\hline
		Learning rate           & 2e-4 \\
		Kernel size             & 5 \\
		Alpha/Beta              & 1 \\
		Batch size              & 1 \\
		Buffer Size             & 1400 \\
		Discriminator repetition  & 1 \\
		Number of Telescope        & 4 \\
		Output Channels         & 1 \\
		Lambda                  & 100 \\
		\hline
	\end{tabular}
\end{table}
This optimized and trained Model is then subjected to testing and evaluation using the Testing Set. The results of this testing and evaluation is presented in the following section.

The Pix2Picx cGAN architecture along with the choice of the values of the hyperparameters mentioned in the Table -\ref{tab:hyperparameters} and used in the implementation of this architecture constitutes the cGAN model (hereinafter referred to as ``the GAN Model'' or simply ``the Model'').

>>>>

In the training of GANs, one often-followed strategy to potentially boost performance is to give the Discriminator an advantage by increasing its number of training epochs before returning to the Generator's training. This hyperparameter is referred to as the Discriminator repetition (as seen in Table \ref{tab:hyperparameters}). While this can lower the Discriminator loss, as shown in Fig.~\ref{fig:Plot_discrep_loss}, it also increases training time. In training our model, we did not notice any significant advantage derived from this strategy. Since the generated images did not noticeably improve with additional Discriminator training, we adopted the strategy of training both the networks with equal preference (Discriminator repetition = 1).

@@

An optimized set of hyperparameters is selected through an iterative process of training and validation. For a tentatively chosen set of the hyperparameters, the model is trained using the Training Set until the both the Discriminator and the Generator loss functions are minimized. This model, thus trained along with its model parameters, is then passed through validation using the Validation Set. This cycle of training and validation is iterated till an optimal set of hypermaraters is arrived at. During each epoch of training of the model, both the Discriminator and the Generator networks are trained for 100,000 steps. Plots of the ''Discriminator Loss'' function and the ``Generator Total Loss'' function presented in Fig.{\ref{fig:Plot_discrep_loss}} and Fig.{\ref{fig:Plot_telescopes_loss} represent the results of this training for two of the hyperparameters, namely, the Disriminator repetition (``discrep'', in short) and the Number of telescopes, respectively}. Obviously, the most compute-intensive part of this process is that of the training of the Model. The results of training and validation presented in this work were carried out on a CPU using two nodes, each with 48 threads and the entire process of training and validation required approximately 20 hours on the machine employed for this work. This iterative process of training and validation of the Model is represented schematically in the Fig.\ref{fig:FlowDiagram}. The chosen optimal set of hyperparameters is presented in the table Table-{\ref{tab:hyperparameters}.}
\begin{table}[ht]
	\centering
	\caption{Selected hyperparameters used for training the model.}
	\label{tab:hyperparameters}
	\begin{tabular}{ll}
		\hline
		\textbf{Hyperparameter} & \textbf{Value} \\
		\hline
		Learning rate           & 2e-4 \\
		Kernel size             & 5 \\
		Alpha/Beta              & 1 \\
		Batch size              & 1 \\
		Buffer Size             & 1400 \\
		Discriminator repetition  & 1 \\
		Number of Telescope        & 4 \\
		Output Channels         & 1 \\
		Lambda                  & 100 \\
		\hline
	\end{tabular}
\end{table}
This optimized and trained Model is then subjected to testing and evaluation using the Testing Set. The results of this testing and evaluation is presented in the following section.

The Pix2Pix cGAN architecture along with the choice of the values of the hyperparameters mentioned in the Table -\ref{tab:hyperparameters} and used in the implementation of this architecture constitutes the cGAN model (hereinafter referred to as ``the GAN Model'' or simply ``the Model'').
